---
title: "R Notebook"
output: html_notebook
---


```{r}
library(magicaxis)
library(pracma)
library(cooltools)
```


Question 1
1.a
```{r}
choose = function(n, k) {
  if (k == 0) {
    return(1)
  } else {
    return(factorial(n) / (factorial(k) * factorial(n - k)))
  }
}

f = function(N) {
  factorial_prod = 1
  j = 0
  
  for (i in 1:N) {
    j = j + i
    factorial_prod = factorial_prod * choose(j, i)
  }
  
  return(factorial_prod)
}

# Test the function
N = 4
distinct_arrangements = f(N)
cat("Number of distinct arrangements for", N, "types:", distinct_arrangements, "\n")

```
```{r}

f2 = function(N) {
  factorial_prod = factorial(sum(0:N))
  
  
  for (i in 1:N) {
    factorial_prod = factorial_prod / factorial(i)
  }
  
  return(factorial_prod)
}

# Test the function
N = 4
distinct_arrangements = f2(N)
cat("Number of distinct arrangements for", N, "types:", distinct_arrangements, "\n")

```
1.b.

```{r}
# Function to calculate ln(f(N)) based on the formula
ln_f_N <- function(N) {
  sum_lfactorials <- sum(lfactorial(2:N))
  ln_f_N <- lfactorial(sum(1:N)) - sum_lfactorials
  return(ln_f_N)
}

# Generate N values from 1 to 10^4
N_values <- 1:1e4

# Calculate ln(f(N)) for each N value
ln_f_values <- sapply(N_values, ln_f_N)

# Plot ln(f(N))
plot(N_values, ln_f_values, type = "l", xlab = "N", ylab = "ln(f(N))",
     main = "Plot of ln(f(N)) for N = 1 to 10^4")

# Add a grid
grid()

```
1.c

```{r}
# Generate N values from 1 to 10^4
N_values <- seq(from = 5*1e4, to = 8*1e4, by = 1e3)

# Calculate ln(f(N)) for each N value
ln_f_values <- sapply(N_values, ln_f_N)

# Plot ln(f(N))
plot(N_values, ln_f_values, type = "l", xlab = "N", ylab = "ln(f(N))",
     main = "Plot of ln(f(N)) for N = 5*1e4, to = 8*1e4")

# Interpolate to find N value corresponding to y=(10**10) / 0.4343
target_y <- (10**10) / 0.4343
interpolated <- approx(ln_f_values, N_values, xout = target_y)

# Add a vertical line to indicate the interpolated N value
abline(v = interpolated$y, col = "red")

# Print the interpolated N value
cat("Interpolated N value:", interpolated$y, "\n")

```


## better way to search?


```{r}
# Target value
target_value <- (10**10) / 0.4342944819

# Set an initial value for N and difference
closest_N <- NULL
closest_diff <- Inf

# Iterate through N values and find the closest one
for (N in 65010:65050) {
  ln_f_value <- ln_f_N(N)
  diff <- abs(ln_f_value - target_value)
  
  if (diff < closest_diff) {
    closest_N <- N
    closest_diff <- diff
  }
}

# Print the result
cat("Integer value of N:", closest_N, "\n")
cat("Closest ln_f_N(N) value:", ln_f_N(closest_N), "\n")

```
Question 2
2.a

```{r}
# Define the n values to plot
n_values_to_plot <- c(2, 3, 5, 10, 15,20,30)

# Initialize an empty list to store the integral values for each n
integral_values_list <- list()

# Perform Monte Carlo integration for each value of n
for (n in n_values_to_plot) {
  # Define the function to integrate
  multisin <- function(x) {
    result <- numeric(n)
    result[1] <- x[1]
  
    for (i in 2:n) {
      result[i] <- sin(result[i - 1] ** x[i])
    }
  
    return(result[n])
  }
  
  # Define the region of integration [0, 1]^n
  lower_limits <- rep(0, n)
  upper_limits <- rep(1, n)
  
  # Perform Monte Carlo integration using mcintegral
  result <- mcintegral(multisin, lower_limits, upper_limits, 1e5, qmc = TRUE)
  
  # Store the estimated integral value
  integral_values_list[[as.character(n)]] <- result$value
}

# Create a plot
plot(n_values_to_plot, unlist(integral_values_list), type = "o", pch = 19, 
     xlab = "Dimension (n)", ylab = "Estimated Integral Value",
     main = "Estimated Integral Value for Different Dimension Values",
     ylim = c(0, 1.5), col = "blue")

# Add labels to each point
text(n_values_to_plot, unlist(integral_values_list), labels = round(unlist(integral_values_list), 4), pos = 3, col = "blue")

# Print the value of the last integration result
cat("Value of the last integration result:", round(integral_values_list[[length(n_values_to_plot)]], 4), "\n")

```
We can see that the integral value approaches 0.768 as n grows. I expect the value as n-> infinity to be 0.768

Q2.b

```{r}


# Define the function
f <- function(x, L=L1) {
  exp_term <- exp(-(1/L)*(((x[,1]-L/2)^2+(x[,2]-L/2)^2+(x[,3]-L/2)^2)^0.5 -
                 ((x[,4]-L/2)^2+(x[,5]-L/2)^2+(x[,6]-L/2)^2)^0.5))
  numerator <- (-1/2/sqrt((x[,1]-x[,4])^2+(x[,2]-x[,5])^2+(x[,3]-x[,6])^2))
  return(exp_term * numerator)
}

# Iterate through L1 values from 1 to 10
for (L1 in 1:10) {
  Q <- mcintegral(f, rep(0, 6), rep(L1, 6), 1e5, qmc = TRUE)
  cat(sprintf('For L1 = %d, Q = %.4f +/- %.4f\n', L1, Q$value, Q$error))
}

# Initialize vectors to store results
L1_values <- 1:10
Q_values <- numeric(length(L1_values))
Q_errors <- numeric(length(L1_values))

# Iterate through L1 values and calculate results
for (i in 1:length(L1_values)) {
  Q <- mcintegral(f, rep(0, 6), rep(L1_values[i], 6), 1e5, qmc = TRUE)
  Q_values[i] <- Q$value
  Q_errors[i] <- Q$error
}

# Create a data frame for plotting
results_df <- data.frame(L1 = L1_values, Q = Q_values, Q_error = Q_errors)

# Plot the results
ggplot(results_df, aes(x = L1, y = Q)) +
  geom_line() +
  geom_errorbar(aes(ymin = Q - Q_error, ymax = Q + Q_error), width = 0.2) +
  labs(title = "Gravitational Mass Fraction vs. L1",
       x = "L1", y = "Gravitational Mass Fraction (Q)") +
  theme_minimal()
```
### Add log log . but clearly is ^5 wrt to L. This also was inline with intuition, as example in lecture had difference of the 1/L^6 in int. Taken L=1 term as the numeric result.

## G factored out as in lecture example and rho_0 factored out from the m1*m2 density matrix 


```{r}
cat("Gravitational binding energy",Q_values[1],"* G * rho_0^2 * L^5" )

```


Question 3

```{r}
library(stats)


# Define the function g.freeman
g.freeman = function(x) {
  q = x / 2
  x * sqrt((besselI(q, 0) * besselK(q, 0) - besselI(q, 1) * besselK(q, 1)) / 2)
}

# Find the radius value rmax where circular velocity is maximal
rmax = optimize(g.freeman, interval = c(0, 5), maximum = TRUE)$maximum

mass_frac = 1 - (rmax + 1) * exp(-rmax)

# Print the value of rmax and mass fraction
cat("Radius value rmax (r/R) where circular velocity is maximal:", rmax, "\n")
cat("Mass Fraction contained within rmax:", mass_frac, "\n")

```

From Lecture notes mass fraction rad/R<x = 1 - (x+1)*e^(-x)
Code for g was taken from lecture notes. 


Question 4

```{r}
# Define the function
f <- function(x) {
  return((x[1]^2 + x[2] - 3)^2 + (x[1] + x[2]^2 - 5)^2)
}

# Run SANN optimization with multiple random initial guesses
num_guesses <- 30
minima_found <- list()

for (i in 1:num_guesses) {
  initial_guess <- runif(2, -10, 10)  # Random initial guess
  result <- optim(initial_guess, f, method = "SANN")
  
  unique_minima <- TRUE
  if (length(minima_found) > 0) {
    for (j in 1:length(minima_found)) {
      distance <- sqrt(sum((result$par - minima_found[[j]]$point)^2))
      if (distance < 0.05) {
        unique_minima <- FALSE
        break
      }
    }
  }
  
  if (unique_minima) {
    minima_found[[length(minima_found) + 1]] <- list(point = result$par, value = result$value)
  }
}

# Display the found unique minima
for (i in 1:length(minima_found)) {
  cat("Local minimum", i, "\n")
  cat("Optimal point:", minima_found[[i]]$point, "\n")
  cat("Optimal value:", minima_found[[i]]$value, "\n\n")
}

```

```{r}
# Define the function
f <- function(x) {
  return((x[1]^2 + x[2] - 3)^2 + (x[1] + x[2]^2 - 5)^2)
}

# Generate x and y values
x <- seq(-4, 4, length.out = 500)
y <- seq(-4, 4, length.out = 500)

# Create a grid of x and y values
grid <- expand.grid(x = x, y = y)

# Calculate the function values for each point in the grid
values <- apply(grid, 1, f)

# Reshape the values to a matrix for contour plotting
z <- matrix(values, nrow = length(x))

# Create the contour plot
contour(x, y, z, nlevels = 100, main = "Contour Plot of f(x, y)", xlab = "x", ylab = "y")

# Add the found minima points to the plot
for (i in 1:length(minima_found)) {
  points(minima_found[[i]]$point[1], minima_found[[i]]$point[2], col = "red", pch = 3)
}

```

I also ran the code to find minima on the negative of the function, it did not return any local maxima, i.e. only returned values very close to the boundary of the search (x or y approaching 500)





